### 2012-06-20 trunk/ijcis-pos-model/sources/abstract.tex
final draft

%%%%%%%%%%%%%%%%%%%%%%%%
%
% ABSTRACT
%
%%%%%%%%%%%%%%%%%%%%%%%%
In real world, it is very common that some objects or concepts have properties with a time-variant or time-related nature. Modelling this kind of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database. Therefore, temporal database models have been proposed to deal with this problem in the literature. Time can be affected by imprecision, vagueness and / or  uncertainty, since existing time measuring devices are inherently imperfect. Additionally, human beings manage time using temporal indications and temporal notions, which may also be imprecise. However, the imperfection in human-used temporal indications is supported by human interpretation, whereas information systems need appropriate support in order to accomplish this task. Several proposals for dealing with such imperfections when modelling temporal data exist. Some of these proposals transform the temporal data into a compact representation. There is not a model for managing and handling uncertainty regarding temporal information.
In this work we present a novel model to deal with imprecision in valid-time databases together with the data manipulation language, \emph{DML} is defined and implemented.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%
%
% ABSTRACT
%
%%%%%%%%%%%%%%%%%%%%%%%%
In real world, it is very common that some objects or concepts have properties with a time-variant or time-related nature. Modelling this kind of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database and have to be appropriately managed. Therefore, temporal database models have been proposed to deal with this problem in the literature. Time can be affected by imprecision, vagueness and / or  uncertainty, since existing time measuring devices are inherently imperfect. Additionally, human beings manage time using temporal indications and temporal notions, which may also be imprecise. However, the imperfection in human-used temporal indications is supported by human interpretation, whereas information systems need appropriate support in order to accomplish this task. Several proposals for dealing with such imperfections when modelling temporal data exist. Some of these proposals transform the temporal data into a compact representation but there is not a formal model for managing and handling uncertainty regarding temporal information.
In this work we present a novel model to deal with imprecision in valid-time databases together with the definition and implementation of the data manipulation language, \emph{DML}.

### 2012-06-20 trunk/ijcis-pos-model/sources/conclusions.tex
final draft

We presented a valid-time model to represent and handle ill-known temporal intervals. The semantics and the implementation of the DML operations are described within this work. As future work , the definition of the Data Definiton Language as well as the querying will be considered.

>>>>

In this work, we presented a complete valid-time model to represent and handle ill-known temporal intervals. The paper includes the formal definition of possibilistic valid-time period in order to manage the time and the formal definition of ill-known constraints to define operators and integrity. This is the first formal model in the literature for possibilistic valid-time in relational databases. The semantics and the implementation of the DML operations are described within this work. As future work , the definition of the Data Definition Language as well as the querying will be considered.

### 2012-06-20 trunk/ijcis-pos-model/sources/intro.tex
version sent to Olga and Nicolas
final draft

The concept of time is very complex to handle and interpret~\cite{Klein1994,Shackle1961}, though it is very natural and omnipresent in real world data. As information systems attempt the modelling of natural objects, concepts or processes, they often require modelling temporal aspects or concepts. Thus, several proposals have arose to obtain theoretical models that allow the modelling or representation of time~\cite{Bolour1982,VanderCruyssen1997}.

A very specific type of information systems are database systems. A database contains data representing real objects or concepts. In reality, some aspects or properties of objects or concepts are time-variant or time-related. e.g., the moment of a bank transaction is traditionally a point in time and thus a time-related notion, the function of an employee in a company can change through the recorded history and is thus time-variant. A temporal database schema is a database schema that models objects or concepts with time-related or time-variant properties. However, the modelling of temporal aspects has a direct impact on the consistency of the temporal database, because the temporal nature of these aspects imposes extra integrity constraints. An example.

Consider a hospital database containing data about the state of the patients in the hospital. Two dates are stored: the date when the patient arrives to and leaves from the hospital. It is clear that a patient cannot left the hospital if he or she has not arrived. Without further precautions, a hospital employee could insert a patient who is already in the hospital. A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.

@@

Notably, in 1994, ``A Consensus Glossary of Temporal Database Concepts'' was published~\cite{Dyreson1994}. For this publication, 44 temporal database researchers, among these some of the main researchers in this field, cooperated to reach a consensus on the nature and definitions of some of the main temporal database concepts and terminology. This glossary was updated in 1998~\cite{Dyreson1998}.

@@

An interesting issue in temporal modelling concerns relationships between temporal notions. In this sense, Allen~\cite{Allen1983} studied temporal relationships between time intervals (and as a special case time points). Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allowed richer and more complex user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query. For example, a query like `who were the department heads when Thomas worked for the institution' can be evaluated using operators similar to Allen's ones.

Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}, and they often have to deal with imperfections like imprecision, vagueness, uncertainty or inconsistencies possibly contained in the descriptions of these temporal notions. Among many others, these possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling. Consider as an example the description of the temporal notion in a sentence like `The Belfry of Bruges was finished between 01/01/1201 A.D. and 31/12/1300 A.D.' contains imperfection because of the uncertainty in the used time-related expression. It is known that the building was finished on a single day, but it is not precisely known  which day this was.

To allow information systems to cope with these and similar data imperfections, many approaches adopt fuzzy sets~\cite{Zadeh1965} for the representation of temporal information~\cite{Mitra1994,Nagypal2003,Billiet2011,Dubois2003}. The temporal relationships studied by Allen were fuzzified by several authors~\cite{Ohlbach2004,Nagypal2003,Schockaert2008}. Although,  Garrido et al. ~\cite{Garrido2009} present a compact representation for the time, they defined different relationships among time intervals by using a combination of regular fuzzy comparisons. Also ~\cite{Garrido2009,Pons2011} studied uncertainty in temporal expressions concerning time intervals. Other approaches, like~\cite{Qiang2009}, use rough sets~\cite{Pawlak1995} to represent time intervals.

@@

In addition to temporal modelling, some attention has been spent on temporal reasoning~\cite{Allen1983}. Although temporal reasoning is not discussed in this paper, it should be noted that, among others, Dubois and Prade et al.~\cite{Dubois2003,DuBois1989} have dealt with fuzziness and uncertainty in temporal reasoning.

The present work defines and implement a model for properly represent and manage uncertainty in valid-time specification in a relational database. None of the previous research offer a database model for acomplish this task.

The aim of this work is to present and explain the main operations of the manipulation language for a temporal database which stores the valid-time periods of the objects affected by imprecision. The rest of the work is organized as follows. Section \ref{sec:prelim} presents some background concepts about both possibility theory and temporal databases. In section \ref{sec:time-rep} the representation of the valid-time intervals in the database is explained. Section \ref{sec:temporal-model} explains the main concepts and behaviour of the temporal model and the Data Manipulation Language (DML) is described and implemented. Finally, Section \ref{sec:conclusions} presents the conclusions and some possibilities for future research work.

>>>>

The concept of time is very complex to handle and interpret~\cite{Klein1994,Shackle1961}, though it is very natural and omnipresent in real world data. As information systems attempt the modelling of natural objects, concepts or processes, they often require modelling temporal aspects or concepts. Thus, several proposals have arisen to obtain theoretical models that allow the modelling or representation of time~\cite{Bolour1982,VanderCruyssen1997}.

A very specific type of information systems are database systems. A database contains data representing real objects or concepts. In real world, some aspects or properties of objects are time-variant or time-related. e.g., the moment of a bank transaction and the status of an employee in a company, are time-related and time-variant notions, respectively.
A temporal database schema is a database schema that models objects with time-related or time-variant properties. However, the modelling of temporal aspects has a direct impact on the consistency of the temporal database, because the temporal nature of these aspects imposes extra integrity constraints and suitable ways of interactive with the human user.

For example, let us consider a hospital database containing data about the state of the patients in a given hospital. Two dates are stored: the date when the patient arrives to and leaves from the hospital. It is clear that a patient cannot leave the hospital if he or she has not arrived. Without further cautions, a hospital employee could insert a patient who is already in the hospital. A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.

@@

Notably, in 1994, ``A Consensus Glossary of Temporal Database Concepts'' was published~\cite{Dyreson1994}. For this publication, 44 temporal database researchers, among them, some of the main researchers in this field, cooperated to reach a consensus on the nature and definitions of some of the main temporal database concepts and terminology. This glossary was updated in 1998~\cite{Dyreson1998}.

@@

An interesting issue in temporal modelling concerns relationships between temporal notions. In this sense, Allen~\cite{Allen1983} studied temporal relationships between time intervals (and as a special case time points). Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allowed more powerful user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query. For example, a query like `who were the department heads when Thomas worked for the institution' can be evaluated using operators similar to Allen's ones.

Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}, and they often have to deal with imperfections like imprecision, vagueness, uncertainty or inconsistencies possibly contained in the descriptions of these temporal notions. These possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling. Consider as an example the description of the temporal notion in a sentence like `The Belfry of Bruges was finished between 01/01/1201 A.D. and 31/12/1300 A.D.'. These sentence contains imperfection because of the uncertainty in the used time-related expression. It is known that the building was finished on a single day, but this day is not precisely known.

To allow information systems to cope with these and similar data imperfections, many approaches adopt fuzzy sets~\cite{Zadeh1965} for the representation and management of temporal information~\cite{Mitra1994,Nagypal2003,Billiet2011,Dubois2003}. The temporal relationships studied by Allen were fuzzified by several authors~\cite{Ohlbach2004,Nagypal2003,Schockaert2008}.   Garrido et al. ~\cite{Garrido2009} presented a compact representation for the time and defined different relationships among time intervals by using a combination of regular fuzzy comparisons. Also ~\cite{Garrido2009,Pons2011} studied uncertainty in temporal expressions concerning time intervals. Other approaches, like~\cite{Qiang2009}, use rough sets~\cite{Pawlak1995} to represent time intervals.

@@

In addition to temporal modelling, some attention has been paid to temporal reasoning~\cite{Allen1983}. Although temporal reasoning is not discussed in this paper, it should be noted that, among others, Dubois and Prade et al.~\cite{Dubois2003,DuBois1989} have dealt with fuzziness and uncertainty in temporal reasoning.

The present work defines and implement a model for properly represent and manage uncertainty in valid-time specification in a relational database. Our work is focused on both the proposal of an appropriate formal framework to suitably manage time in databases and the implementation of an DML that allows to the user the transparent use of this proposal. None of the previous research offer a database model to accomplish this task.

This way together with the theoretical model, we also present and explain the main operations of the manipulation language for a temporal database which stores the valid-time periods of the objects affected by imprecision. The rest of the work is organized as follows. Section \ref{sec:prelim} presents some background concepts about both possibility theory and temporal databases. In section \ref{sec:time-rep} the representation of the valid-time intervals in the database is explained. Section \ref{sec:temporal-model} explains the main concepts of the temporal Data Manipulation Language (DML) and its implementation. Finally, Section \ref{sec:conclusions} presents the conclusions and some guidelines for future research work.

### 2012-06-20 trunk/ijcis-pos-model/sources/preliminaries.tex
final draft

\begin{definition}
Consider a possibility measure $\Pi$ on $\Omega$. A \emph{possibility distribution} $\pi$ on $\Omega$ underlying the possibility measure $\Pi$ is then a function defined by:
	\begin{align}
	\pi : \Omega \rightarrow \left[0, 1\right] : \pi(u) = \Pi(\{u\})
	\end{align}
\end{definition}

\begin{definition}
Consider a confidence measure $N$ on a set of outcomes $\Omega$. Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$. $N$ is now a \emph{necessity measure} on $\Omega$ if it satisfies:
	\begin{align}
	N\left(\bigcap_{j \in J} A_{j} \right) = \inf_{j \in J} N(A_{j})
	\end{align}
\end{definition}

@@

It is important to understand the difference between the following two concepts:
\begin{itemize}
\item
A \emph{possibilistic variable} $X$ is bounded to take only one value , but this value is not known due to incomplete knowledge. 
\item
An \emph{ill-known set}~\cite{Dubois1988a}: a possibilistic variable defined over the universe $\Pow(U)$.
\end{itemize}

@@

This work deals with ill-known intervals. These are ill-known sets, defined and represented via a starting and ending points, which are ill-known values. The elements of the set are the values between the starting and ending points. A closed ill-known interval with starting point defined by possibilistic variable $X$ and ending point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$. %The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.

@@

\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=0.25]{./graphs/triangular.pdf}

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:triangular-dist}Example of fuzzy number.}
%  \label{fig:fuzzy-validity-period}
\vspace*{13pt}

If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}. 
%In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number (Figure \ref{fig:triangular-dist}).

@@

\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=0.25]{./graphs/trapezoidalDistribution.pdf}

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:trapezoidal-dist}Example of fuzzy interval.}
%  \label{fig:fuzzy-validity-period}
\vspace*{13pt}
%\def\JPicScale{0.5}
%\begin{figure}[h!]
%  \centering
%  \input{./graphs/trapezoidal.tex}
%  \caption{Trapezoidal membership function}
%  \label{fig:trapezoidal}
%\end{figure}

@@

\begin{definition}
Given a universe $U$, an ill-known constraint $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value denoted by its possibilistic variable $V$ over $U$, i.e.:
\begin{align}
\label{eq:ill-known-constraint}
C \triangleq (R,V)
\end{align}
Set $A$ satisfies the constraint if and only if:
\begin{align}
\forall a \in A : (a,V) \in R
\end{align}
\end{definition}

@@

\begin{example}
As example, consider  $X = \left[5, 3, 2 \right]$ and let $C =(\leq, X)$ the ill-known constraint. Then, the evaluation of the possibility and the necessity are obtained from \eqref{ill-known-pos} and \eqref{ill-known-nec} respectively. (See Figure \ref{fig:example-ill-known}).

\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=1]{./graphs/example-ill-known.pdf}
}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:example-ill-known}Example of the evaluation of the ill-known constraint $C \triangleq (\leq, X)$. Possibility and necessity measures are shown in grey. }
%  \label{fig:fuzzy-validity-period}
\vspace*{13pt}

\begin{align}
\Pos(A\text{ satisfies }C) & =\\
\nonumber
\min_{a \in A}\left(\sup_{a \leq R}\pi_{X}(w)\right)\\
\Nec(A\text{ satisfies }C) & =\\
\nonumber
\min_{a \in A}\left(\inf_{a > w} 1-\pi_{X}(w)\right) 
\end{align}

@@

It is observed that Boolean combinations of constraints are required. For example, the problem of interval evaluation (explained earlier) requires that all elements of an interval $[a,b]$ are larger than a value $X$ and, at the same time, smaller than a value $Y$, which implies that a conjunctive Boolean combination of both constraints must be satisfied. To allow Boolean combinations of constraints, the following definitions are introduced.
\begin{definition}
\label{def:set-evaluation}
Consider a universe $U$, an $n$-ary vector $\mathbf{C}$ of constraints and a Boolean function $\bool:\mathbb{B}^{n}\rightarrow\mathbb{B}$. An evaluation function is defined by:
\begin{equation}
\lambda:\Pow(U)\rightarrow\mathbb{B}:\lambda(A)\mapsto\bool\Big(C_{1}(A),...,C_{n}(A)\Big).
\end{equation}
\end{definition}
Definition \ref{def:set-evaluation} presents the definition of an evaluation function that evaluates a Boolean combination of some basic constraints. Informally, it states that a set $A$ passes the evaluation made by $\lambda$ if the Boolean combination of some propositions equals $T$. This crisp definition can be generalized to the case of ill-known constraints.
\begin{definition}
\label{def:ill-known-sets}
Consider a universe $U$, an $n$-ary vector $\mathbf{C}$ of ill-known constraints and a Boolean function $\bool:\mathbb{B}^{n}\rightarrow\mathbb{B}$. The uncertainty about the evaluation of a set $A$ by an evaluation function $\lambda$ is then given by:
\begin{equation}
\forall A\in\Pow(U):\pi_{\lambda(A)}=\widetilde{\bool}\Big(\pi_{C_1(A)},...,\pi_{C_n(A)}\Big)\\
\end{equation}
Hereby, $\widetilde{\bool}$ is the possibilistic extension of $\bool$.
\end{definition}
It is well known that any Boolean function $\bool$ can be cast to a canonical form~\cite{McCluskey1965}, requiring only logical conjunction $\wedge$, logical disjunction $\vee$ and logical negation. Therefore, only those connectives will be treated within the scope of this paper. By applying the possibilistic extensions of $\wedge$, $\vee$ and $\neg$, concrete equations are obtained for the calculations of uncertainty about the evaluation of a set by means of an evaluation function $\lambda$. In the case of conjunction (i.e., $\bool=\wedge$), the inference of uncertainty about the evaluation of a set reduces to:
\begin{eqnarray}
\label{eq:conjunctive1}
\forall A\in\Pow(U):\Pos(\lambda(A))&=\\
\nonumber
\min_{i=1 \ldots n}\Pos\left(C_i(A)\right)\\
\label{eq:conjunctive2}
\forall A\in\Pow(U):\Nec(\lambda(A))&=\\
\nonumber
\min_{i=1 \ldots n}\Nec\left(C_i(A)\right).
\end{eqnarray}
In the case of disjunction (i.e. $\bool=\vee$), the inference of uncertainty about the evaluation of a set reduces to:
\begin{eqnarray}
\label{eq:disjunctive}
\forall A\in\Pow(U):\Pos(\lambda(A))&=\\
\nonumber
\max_{i=1 \ldots n}\Pos\left(C_i(A)\right)\\
\forall A\in\Pow(U):\Nec(\lambda(A))&=\\
\nonumber
\max_{i=1 \ldots n}\Nec\left(C_i(A)\right).
\end{eqnarray}
Note that by using the functions $\min$ and $\max$ here, there is an implicit assumption that the possibilistic variables $\pi_{C_i}$ are mutual $\min$-dependent in the sense of De Cooman (i.e. non-interactive). For an extensive reading on (in)dependency of possibilistic variables, the reader is referred to~\cite{GertDeCooman1997b},\cite{GertDeCooman1997a},\cite{GertDeCooman1997}. In case of $\neg$, we get:
\begin{eqnarray}
\label{eq:negation}
\forall A\in\Pow(U):\Pos(\neg\lambda(A))&=\\
\nonumber
1-\Nec(\lambda(A))\\
\forall A\in\Pow(U):\Nec(\neg\lambda(A))&=\\
\nonumber
1-\Pos(\lambda(A)).
\end{eqnarray}

\begin{example}
Consider that we want to check if crisp interval $I = \left[j, k\right]$ is included in $\left[X, Y\right]$. In this situation, two ill-known constraints are constructed.

@@

\begin{eqnarray}
C_1 & \triangleq\left(\geq,X\right)\\
C_2 & \triangleq\left(\leq,Y\right)
\end{eqnarray}

@@

There is a special boolean combination of constraints that is of particular interest; let us see it.
\begin{definition}
\emph{Conjunctive combination of ill-known constraints}. Consider a universe $U$. Let $R_1, R_2 \subseteq U$ be two binary relations in U. Let $X_1, X_2$ be two fixed ill-known values in $U$. Let  $C_1 = (R_1, X_1)$ and $C_2 = (R_2, X_2)$ be two ill-known constraints. A conjunctive combination of both constraints is given by:
\begin{align}
\label{eq:convex-combination}
CC \triangleq \left \lbrace C_1 \wedge C_2 \right \rbrace
\end{align}
In a more general way, it is possible to define the conjunctive combination of an n-ary vector of constraints:
\begin{align}
\label{eq:nary-convex-combination}
CC \triangleq \left \lbrace C_1 \wedge \ldots \wedge C_n \right \rbrace
\end{align}
\end{definition}

@@

\begin{definition}
\label{def:convex-combination-ill-known-constraints}
\emph{Ill-known value defined by conjunctive combination of constraints.}
Consider a universe $U$, $X$ an ill-known value and $CC \triangleq \left \lbrace C_{1_{Z_1}} \wedge \ldots \wedge C_{n_{Z_n}}  \right \rbrace$ a convex combination of ill-known constraints over the variables $Z_1, \ldots, Z_n$. Then, the value $X$ is defined by means of the convex combination:
\begin{align}
\label{eq:ill-known-value-by-convex-constraints}
X \triangleq CC 
\end{align}

The uncertainty about the evaluation of an ill-known value $X$ is given by:
\begin{align}
\label{eq:ill-known-value-def-by-const-unc}
X \in \Pow(U):\pi_{X}=\pi_{CC}
\end{align} 
Note that $\pi_{X}$ is convex since $\pi_{CC}$ is convex as demonstrated in Theorem \ref{th:convex-combination-ill-known-constraints}.
\end{definition}

@@

\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=1]{./graphs/ill-known-by-constraints.pdf}

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:example-ill-known-by-const}Ill-known values $X$ and $Y$. The grey area represents the ill-known value $Z$ defined by the convex combination of the two ill-known constraints $C_1$ and $C_2$.}
\vspace*{13pt}

@@

%enlace:
We have seen the main theoretical concepts about ill-known values. Now, we are going to explain the main concepts about the treatment of time and the imperfection related to the time in databases. In the following subsections we will show how to represent imperfect time intervals by means of ill-known values.

>>>>

\begin{definition}
Consider a possibility measure $\Pi$ on $\Omega$. A \emph{possibility distribution} $\pi$ on $\Omega$ underlying the possibility measure $\Pi$ is a function defined by:
	\begin{align}
	\pi : \Omega \rightarrow \left[0, 1\right] : \pi(u) = \Pi(\{u\})
	\end{align}
\end{definition}

\begin{definition}
Consider a confidence measure $N$ on $\Omega$. Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$. $N$ is a \emph{necessity measure} on $\Omega$ if it satisfies:
	\begin{align}
	N\left(\bigcap_{j \in J} A_{j} \right) = \inf_{j \in J} N(A_{j})
	\end{align}
\end{definition}

@@

% It is important to understand the difference between the following two concepts:
% \begin{itemize}
% \item
% A \emph{possibilistic variable} $X$ is bounded to take only one value , but this value is not known due to incomplete knowledge. 
% \item
% An \emph{ill-known set}~\cite{Dubois1988a}: a possibilistic variable defined over the universe $\Pow(U)$.
% \end{itemize}

@@

This work deals with ill-known intervals. These are ill-known sets, defined and represented via a starting and ending points that, in turn are ill-known values. The elements of the set are the values between the starting and ending points. A closed ill-known interval with starting point defined by possibilistic variable $X$ and ending point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$. %The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.

@@

\begin{samepage}
\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=0.25]{./graphs/triangular.pdf}

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:triangular-dist}Example of fuzzy number.}
%  \label{fig:fuzzy-validity-period}
\vspace*{13pt}
\end{samepage}
If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}. 
%In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number (Figure \ref{fig:triangular-dist}).

@@

\begin{samepage}
\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=0.25]{./graphs/trapezoidalDistribution.pdf}

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:trapezoidal-dist}Example of fuzzy interval.}
%  \label{fig:fuzzy-validity-period}
\vspace*{13pt}
\end{samepage}

%\def\JPicScale{0.5}
%\begin{figure}[h!]
%  \centering
%  \input{./graphs/trapezoidal.tex}
%  \caption{Trapezoidal membership function}
%  \label{fig:trapezoidal}
%\end{figure}

@@

\begin{definition}
Given a universe $U$, an ill-known constraint $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed ill-known value denoted by its possibilistic variable $V$ over $U$, i.e.:
\begin{align}
\label{eq:ill-known-constraint}
C \triangleq (R,V)
\end{align}
Set $A$ satisfies the constraint if and only if:
\begin{align}
\forall a \in A : (a,V) \in R
\end{align}
\end{definition}

@@

\begin{samepage}
\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=1]{./graphs/example-ill-known.pdf}
}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:example-ill-known}Example of the evaluation of the ill-known constraint $C \triangleq (\leq, X), X = \left[5, 3, 2\right]$. Possibility and necessity measures are shown in grey. }
%  \label{fig:fuzzy-validity-period}
\vspace*{13pt}
\end{samepage}

\begin{example}
Consider  $X = \left[5, 3, 2 \right]$ and let $C =(\leq, X)$ the ill-known constraint. Then, the evaluation of the possibility and the necessity are obtained from \eqref{ill-known-pos} and \eqref{ill-known-nec} respectively. (See Figure \ref{fig:example-ill-known}).

\begin{align}
\Pos(A\text{ satisfies }C) & =\\
\nonumber
\min_{a \in A}\left(\sup_{a \leq w}\pi_{X}(w)\right)\\
\Nec(A\text{ satisfies }C) & =\\
\nonumber
\min_{a \in A}\left(\inf_{a > w} 1-\pi_{X}(w)\right) 
\end{align}

@@

It is observed that Boolean combinations of constraints are required. For example, the problem of interval evaluation (explained earlier) requires that all the elements of an interval $[a,b]$ are larger than a value $X$ and, at the same time, smaller than a value $Y$, which implies that a conjunctive Boolean combination of both constraints must be satisfied. To allow Boolean combinations of constraints, the following definitions are introduced.
\begin{definition}
\label{def:set-evaluation}
Consider a universe $U$, an $n$-ary vector $\mathbf{C}$ of constraints and a Boolean function $\bool:\mathbb{B}^{n}\rightarrow\mathbb{B}$. An evaluation function is defined by:
\begin{equation}
\lambda:\Pow(U)\rightarrow\mathbb{B}:\lambda(A)\mapsto\bool\Big(C_{1}(A),...,C_{n}(A)\Big).
\end{equation}
\end{definition}
Definition \ref{def:set-evaluation} presents a function that evaluates a Boolean combination of some basic constraints. Informally, it states that a set $A$ passes the evaluation made by $\lambda$ if the Boolean combination of some propositions equals $T$. This crisp definition can be generalized to the case of ill-known constraints.
\begin{definition}
\label{def:ill-known-sets}
Consider a universe $U$, an $n$-ary vector $\mathbf{C}$ of ill-known constraints and a Boolean function $\bool:\mathbb{B}^{n}\rightarrow\mathbb{B}$. The uncertainty about the evaluation of a set $A$ by an evaluation function $\lambda$ is then given by:
\begin{equation}
\forall A\in\Pow(U):\pi_{\lambda(A)}=\widetilde{\bool}\Big(\pi_{C_1(A)},...,\pi_{C_n(A)}\Big)\\
\end{equation}
Hereby, $\widetilde{\bool}$ is the possibilistic extension of $\bool$.
\end{definition}
It is well known that any Boolean function $\bool$ can be cast to a canonical form~\cite{McCluskey1965}, requiring only logical conjunction $\wedge$, logical disjunction $\vee$ and logical negation. Therefore, only those connectives will be treated within the scope of this paper. By applying the possibilistic extensions of $\wedge$, $\vee$ and $\neg$, concrete equations are obtained for the calculations of uncertainty about the evaluation of a set by means of an evaluation function $\lambda$. In the case of conjunction (i.e., $\bool=\wedge$), the inference of uncertainty about the evaluation of a set reduces to:
\begin{eqnarray}
\label{eq:conjunctive1}
\forall A\in\Pow(U):\Pos(\lambda(A))&=\\
\nonumber
\min_{i=1 \ldots n}\Pos\left(C_i(A)\right)\\
\label{eq:conjunctive2}
\forall A\in\Pow(U):\Nec(\lambda(A))&=\\
\nonumber
\min_{i=1 \ldots n}\Nec\left(C_i(A)\right).
\end{eqnarray}
In the case of disjunction (i.e. $\bool=\vee$), the inference of uncertainty about the evaluation of a set reduces to:
\begin{eqnarray}
\label{eq:disjunctive}
\forall A\in\Pow(U):\Pos(\lambda(A))&=\\
\nonumber
\max_{i=1 \ldots n}\Pos\left(C_i(A)\right)\\
\forall A\in\Pow(U):\Nec(\lambda(A))&=\\
\nonumber
\max_{i=1 \ldots n}\Nec\left(C_i(A)\right).
\end{eqnarray}
Note that by using the functions $\min$ and $\max$ here, there is an implicit assumption that the possibilistic variables $\pi_{C_i}$ are mutual $\min$-dependent in the sense of De Cooman (i.e. non-interactive). For an extensive reading on (in)dependency of possibilistic variables, the reader is referred to~\cite{GertDeCooman1997b},\cite{GertDeCooman1997a},\cite{GertDeCooman1997}. In case of $\neg$, we get:
\begin{eqnarray}
\label{eq:negation}
\forall A\in\Pow(U):\Pos(\neg\lambda(A))&=\\
\nonumber
1-\Nec(\lambda(A))\\
\forall A\in\Pow(U):\Nec(\neg\lambda(A))&=\\
\nonumber
1-\Pos(\lambda(A)).
\end{eqnarray}

\begin{example}
Consider that we want to check if the crisp interval $I = \left[j, k\right]$ is included in $\left[X, Y\right]$. In this situation, two ill-known constraints are constructed.

@@

\begin{eqnarray}
\label{eq:constraint-c1}
C_1 & \triangleq\left(\geq,X\right)\\
\label{eq:constraint-c2}
C_2 & \triangleq\left(\leq,Y\right)
\end{eqnarray}

@@

There is a special boolean combination of constraints that is of particular interest; let us see it.
\begin{definition}
\emph{Conjunctive combination of ill-known constraints}. Consider a universe $U$. Let $R_1, R_2$ be two binary relations in U. Let $X_1, X_2$ be two fixed ill-known values in $U$. Let  $C_1 = (R_1, X_1)$ and $C_2 = (R_2, X_2)$ be two ill-known constraints. A conjunctive combination of both constraints is given by:
\begin{align}
\label{eq:convex-combination}
CC \triangleq \left \lbrace C_1 \wedge C_2 \right \rbrace
\end{align}
In a more general way, it is possible to define the conjunctive combination of an n-ary vector of constraints:
\begin{align}
\label{eq:nary-convex-combination}
CC \triangleq \left \lbrace C_1 \wedge \ldots \wedge C_n \right \rbrace
\end{align}
\end{definition}

@@

\begin{definition}
\label{def:convex-combination-ill-known-constraints}
\emph{Ill-known value defined by conjunctive combination of constraints.}
Consider a universe $U$, $CC \triangleq \left \lbrace C_{1_{Z_1}} \wedge \ldots \wedge C_{n_{Z_n}}  \right \rbrace$ a conjunctive combination of ill-known constraints over the variables $Z_1, \ldots, Z_n$. The uncertainty about the evaluation of an ill-known value $X$ is given by:
\begin{align}
\label{eq:ill-known-value-def-by-const-unc}
X \in \Pow(U):\pi_{X}=\pi_{CC}
\end{align}

The definition of the ill-known value $X$ with respect to the conjunctive combination of ill-known constraints, is written as:
\begin{align}
\label{eq:ill-known-value-by-convex-constraints}
X \triangleq CC 
\end{align}

Note that $\pi_{X}$ is convex since $\pi_{CC}$ is convex as demonstrated in Theorem \ref{th:convex-combination-ill-known-constraints}.
\end{definition}

@@

\begin{samepage}
\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=1]{./graphs/ill-known-by-constraints.pdf}

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:example-ill-known-by-const}Ill-known values $X$ and $Y$. The grey area represents the ill-known value $Z$ defined by the convex combination of the two ill-known constraints $C_1$ and $C_2$.}
\vspace*{13pt}
\end{samepage}

@@

%enlace:
We have seen the main theoretical concepts about ill-known values. Now, we are going to explain the main concepts about the treatment of time and the imperfection related to the time in databases. This two preliminary analysis will be the pillars of our proposal in section \ref{sec:time-rep}.

%In the following subsections we will show how to represent imperfect time intervals by means of ill-known values.
